\chapter*{Acknowledgements}
\markboth{Acknowledgements}{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}


While my name may be displayed on the title page as the author of this work, none of this would have
been possible without the combined efforts of many others. First and foremost, I want to thank my
supervisors, Yves Revaz and Anne Verhamme, for offering me the opportunity to work on a doctorate
with them. They were nothing but kind and supportive throughout these past four years, and generous
with advice and encouragement. I also want to thank Romain Teyssier, who (aside from innumerable
teachings) referred me to this position, and who took plenty of time to show me how to write a good
paper. Many thanks also to Jean-Paul Kneib, whose support and efforts allowed the position I filled
to exist, and to my thesis defense jury, who in alphabetical order by last name were Michaela
Hirschmann, Joki Rosdahl, and Joachim Stadel, for their many helpful suggestions on how to improve
this manuscript. A very special thank you goes out to Yves Revaz and Joki Rosdahl for their
exceptionally detailed reviews of the first submission of this thesis. \\


The past four years would have been really tough without good coworkers, and I was very lucky to
have excellent ones, which I am happy to call friends today. Lo\"ic Hausammann, who was my senior
by one year and some change, was blessed by the gods with infinite knowledge and had an answer to
life, the Universe, and everything. Whether the answers were correct though was no concern of his;
as long as he had them, things were good. Admittedly, they were quite funny and entertaining though.
He also showed me around the bowels of \swift as I made my first steps into the unknown, and was
eager to help wherever he could. Florian Cabot, who is nothing short of a computational science and
visualization wizard, had the misfortune of sharing an office with me for years. (I'm still not sure
how they managed to stick the only two metalhead vegetarians at the Obs in the same corner office,
but I'm glad they did!). It was great having somebody around with many common interests, and hunting
down ridiculous bugs in our codes was much more fun while overdramatically complaining about them to
each other. And Mahsa Sanati, who has the incredible superpower to make friends in every city she
visits. Unfortunately our individual projects diverged from each other, but she never failed to
infect the room with cheerfulness. Especially when ice cream was involved. Or even mentioned. Thank
you all for all the good times we shared and all the support you offered. It was a pleasure and a
privilege to work with you. \\

I would also like to thank the fantastic people around \swift, in particular to Matthieu Schaller
and Bert Vandenbroucke, for the warm welcome they extended to me when I first started; for the
unending eagerness to offer advice, support, ideas, and help; for the review of what is probably
hundreds of merge requests. Many thanks also to Pedro Gonnet, Peter Draper, Richard Bower, Josh
Borrow, and Jacob Kegerreis. I hope we can continue working together on \swift for many years to
come. \\

My undying gratitude and love is extended to my mother and brother. I do not have the words to
express how grateful I am for all the sacrifices you took upon yourself so I could walk down this
path, so forgive me for keeping it short. The same goes to my dear dear friends, Tony Inchiparambil, Kay B\"ohringer, and Hermann Nienhaus. Without fail, they were there for me with cheer and love, song and drink, an open ear and a shoulder to lean on, in both the best of times and darkest of times. Without you, I wouldn't be who I am, nor where I am today. We partook. \\


Finally, many thanks to the HPC-EUROPA3 project for awarding me a grant to spend time taking
\GEARRT to the next level in Leiden with Matthieu. This work was carried out under the HPC-EUROPA3
project (project number: 730897) with the support of the European Commission Capacities Area -
Research Infrastructures Initiative.



\bigskip
\bigskip


So long, and thanks for all the fish! \\


\noindent\textit{18. February 2023}
\hfill M.~I.

